---
title: "p8105_hw2_msk2265"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 2


Read and clean the Mr. Trash Wheel sheet:

- specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
- use reasonable variable names
-omit rows that do not include dumpster-specific data (not sure what data is not dumpstir specific)

 The data include a column for the (approximate) number of homes powered. This calculation is described in the Homes powered note, but not applied to every row in the dataset. Update the data to include a new homes_powered variable based on this calculation.
 
```{r}
#decided to load the readxl library in case there are functions used in the background other than jusr readxl

mr_trash_wheel_df = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N586") %>% 
  janitor::clean_names() %>% 
  mutate( 
    homes_powered = (weight_tons*500)/30 #formula based on the note in the excel spreadsheet: Homes Powered - Each ton of trash equates to on average 500 kilowatts of electricity.  An average household will use 30 kilowatts per day.
  )
```


Use a similar process to import, clean, and organize the data for Professor Trash Wheel and Gwynnda, and combine these with the Mr. Trash Wheel dataset to produce a single tidy dataset. To keep track of which Trash Wheel is which, you may need to add an additional variable to all datasets before combining.

```{r}
#loading professor trash wheel data in the same way that Mr. trash wheel was loaded

prof_trash_wheel_df = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M108") %>% 
  janitor::clean_names() %>% 
  mutate( 
    homes_powered = (weight_tons*500)/30 #formula same as it was for Mr. trash wheel
  )

#loading gwynnda trash wheel data in the same way that Mr. trash wheel was loaded

gwynnda_trash_wheel_df = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L157") %>% 
  janitor::clean_names() %>% 
  mutate( 
    homes_powered = (weight_tons*500)/30 #formula same as it was for Mr. trash wheel
  )

```

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in the resulting dataset, and give examples of key variables. 

For available data, what was the total weight of trash collected by Professor Trash Wheel? 

What was the total number of cigarette butts collected by Gwynnda in July of 2021?


